\newpage


\section{Proposal}

\textbf{Research Questions:}
\begin{itemize}

  % === HCI & Player Experience ===
  \item How can HCI-inspired experimental designs be combined with large-scale Reddit data to reveal how emotional design elements (UI aesthetics, narrative cues, reward systems) manifest in player discourse online?  
  \item Can deep learning models detect “affective affordances” in games (e.g., mechanics that reliably trigger frustration, joy, or solidarity), and how are these affordances narrated in Reddit communities?  
  \item What role do interface metaphors and interaction rituals play in shaping how emotions are communicated in asynchronous text forums versus synchronous platforms (e.g., Discord, Twitch)?  

  % === Mass Communication & Digital Publics ===
  \item How do emotions function as resources in “affective publics” (Papacharissi, 2015) around games, and can ML methods identify how outrage, hype, or nostalgia mobilize community participation?  
  \item Can sentiment trajectories be linked to agenda-setting and framing effects in mass communication (e.g., how game journalism and influencer content set emotional baselines for Reddit discourse)?  
  \item Do emotional expressions on Reddit reflect broader cultural anxieties (e.g., blockchain, loot boxes, AI NPCs), and how can computational models trace the diffusion of these concerns across online publics?  

  % === Machine Learning & Deep Learning ===
  \item How effective are transformer-based large language models (LLMs) compared to domain-specific models (e.g., fine-tuned RoBERTa on gaming corpora) in capturing nuanced emotional discourse such as irony, nostalgia, or moral disgust?  
  \item Can temporal deep learning architectures (e.g., LSTMs, Transformers with time embeddings) predict emotional “turning points” in community sentiment following major game events (patches, scandals, releases)?  
  \item To what extent can graph neural networks (GNNs) applied to Reddit reply trees model emotional contagion — i.e., how anger or excitement propagates through social networks?  
  \item How can multimodal deep learning (CLIP, Flamingo, VisualBERT) analyze memes, screenshots, and GIFs shared in subreddits to reveal how visual culture amplifies or reinterprets textual emotions?  

  % === Social Science & Identity ===
  \item How do emotions operate as boundary-making practices (e.g., “true fans” vs. “casuals”), and can ML-driven discourse analysis quantify these boundaries over time?  
  \item Can unsupervised clustering methods identify emergent “emotional subcultures” (e.g., nostalgia-driven players vs. competitive grinders) within larger Reddit communities?  
  \item How does emotional alignment within subreddits relate to social identity theory — and can embedding-based similarity measures capture affective in-group vs. out-group dynamics?  

  % === Cross-Cultural / Multilingual Analysis ===
  \item How do multilingual communities express emotion differently (e.g., irony in English, intensifiers in Spanish, symbolic memes in Chinese), and how robust are multilingual LLMs (XLM-R, mBERT) in detecting these cultural variations?  
  \item To what extent does machine translation flatten emotional nuance, and could fine-tuning LLMs on parallel corpora of Reddit multilingual discourse preserve culturally specific affect?  

  % === Predictive & Applied Angles ===
  \item Can hybrid models (deep learning + social network analysis) predict which subreddit threads are likely to become toxic spirals or solidarity hubs?  
  \item Could reinforcement learning simulations model “what if” interventions (e.g., alternative patch note communication, moderator interventions) to forecast emotional outcomes?  
  \item How can explainable AI (XAI) methods bridge computational prediction with social science theory, helping communication scholars interpret why certain emotions dominate at specific lifecycle stages?  

\end{itemize}
